---
title: "Practical Machine Learning Prediction Assignment"
author: "CRK8"
date: "May 29, 2017"
output: html_document
---

## Instructions

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbbell of 6 participants.

## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data

The training data for this project are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r read, include = T, echo = T, cache = T}
train <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
test <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")
```


The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

## Model Prediction

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. 

### Model Build

A random forest machine learning algorithm is a good first choice when building classification model with many potential predictors. Since processing needs for this type of model can be large, parallel processing is used to speed up the process.

After calling the necessary packages, the datasets are cleaned of fields that have all `NA` test data. Then a data partition is used to create training and cross-validation data sets. Finally fields that shouldn't be used for the prediction model are removed.

```{r setup, include = T, echo = T, cache = T}
library(caret, quietly = T)
library(randomForest, quietly = T)
library(parallel, quietly = T)
library(doParallel, quietly = T)

set.seed(88)

tn <- train[,!apply(is.na(test),2,all)] # gets rid of columns with all NA values in test, cant use those for prediction
tt <- test[,!apply(is.na(test),2,all)] 

inVal <- createDataPartition(y=tn$classe, p=0.7, list=F)
tng <- tn[inVal,]
val <- tn[-inVal,]

tng.f <- tng[,-c(1:7)]

```

The random forest model is created using 10-fold cross validation.

```{r model, include = T, echo = T, cache = T}

registerDoParallel(makeCluster(detectCores()-1))
m1 <- train(classe~.,data=tng.f,method="rf",trControl=trainControl(method = "cv",number=10,allowParallel = T))
stopCluster(makeCluster(detectCores()-1))
registerDoSEQ()

```


### Cross Validation and Expected Out of Sample Error

The model predicts the training data set with 100% accuracy as can be seen below.

```{r cv, include = T, echo = T, cache = T, comment=NA}

table(predict(m1,tng.f),tng.f$classe)

```

When the validation data set is used, the accuracy of the model is 99.3% with the 95% confidence interval being > 99.0%. 

```{r se, include = T, echo = T, cache = T, comment=NA}

confusionMatrix(predict(m1,val),val$classe)

```

### Course Project Prediction Quiz

When the random forest machine learning algorithm is applied to the 20 test cases available, the results yield a perfect 20/20 score.

```{r quiz, include = T, echo = T, cache = T, comment=NA}

predict(m1,tt)

```
